The impact of cosmic variance on simulating weak lensing surveys

### 2014-11-04 Writeup/paper_v3.tex
more abstract tweaking and shortening

\begin{abstract}
Upcoming weak lensing surveys will survey large cosmological volumes
in order to measure the growth of cosmological structure with time and
thereby constrain dark energy.  One major systematic uncertainty in
this process is the calibration of the weak lensing shape distortions,
or shears.  Nearly all upcoming surveys plan to test their shear
estimation algorithms using sophisticated image simulations that
attempt to mimick real data as closely as possible, including
realistic galaxy populations based on high-resolution data from the
Hubble Space Telescope ({\em HST}).  However, the existing datasets
from the {\em HST} cover very small cosmological volumes, such that
cosmic variance could cause the galaxy population in them
to be atypical.  This is particularly the case when selecting galaxy
populations in limited redshift slices, which could be dominated by a
single large overdensity or underdensity.  In that case, the
morphology-density relation could cause changes in the local galaxy
populations that would result in incorrect conclusions to be drawn
about the calibration of shear estimates as a function of redshift.
We directly test this scenario using the COSMOS survey, which is the
largest-area {\em HST} survey to date, and show how the
statistical properties of shape (axis ratio or ellipticity) and
morphological parameters such as \sersicn or the bulge-to-total ratio
are influenced by redshift-dependent cosmic variance in this survey. This
study requires a careful distinction between environment effects from
large-scale structure, which we do not wish to include in simulations,
and general trends in the galaxy population with redshift. We conclude
that this effect is large enough to exceed the systematic error budget
in future surveys, but can be mitigated with careful choice of
training dataset and sufficiently large redshift binning. 
\end{abstract}

>>>>

\begin{abstract}
Upcoming weak lensing surveys will survey large cosmological volumes to measure the growth of cosmological structure with time and
thereby constrain dark energy.  One major systematic uncertainty in
this process is the calibration of the weak lensing shape distortions,
or shears.  Most upcoming surveys plan to test their shear
estimation algorithms using sophisticated image simulations that
include 
realistic galaxy populations based on high-resolution data from the
Hubble Space Telescope ({\em HST}).  However, existing datasets
from the {\em HST} cover very small cosmological volumes, so cosmic variance could cause the galaxy populations in them
to be atypical. A narrow redshift slice from such surveys could be dominated by a single large overdensity or underdensity.  In that case, the
morphology-density relation could alter the local galaxy
populations and yield an incorrect calibration of shear estimates as a function of redshift. We directly test this scenario using the COSMOS survey, the largest-area {\em HST} survey to date, and show how the
statistical distributions of galaxy shapes and
morphological parameters (e.g., S\'{e}rsic $n$)
are influenced by redshift-dependent cosmic variance. 
% We must distinguish between environment effects from large-scale structure, which we do not wish to include in simulations, and general trends in the galaxy population with redshift. 
The typical variation in RMS ellipticity due to environmental effects is 5 per cent for redshift bins of width $\Delta z=0.05$, which could result in uncertain shear calibration at the 1 per cent level. We conclude that the cosmic variance effects are large enough to exceed the systematic error budget of future surveys, but can be mitigated with careful choice of training dataset and sufficiently large redshift binning.
\end{abstract}
